---
title: "PA1_template"
author: "TM"
date: "2023-02-12"
output:

  html_document


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data
Question 1: Code for reading in the dataset and/or processing the data
```{r loading}
suppressMessages(library(dplyr))
suppressMessages(library(lubridate))
suppressMessages(library(ggplot2))
df <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
Question 2: Histogram of the total number of steps taken each day
Question 3: Mean and median number of steps taken each day
```{r pressure}
steps_per_day <- df %>% 
  group_by(date) %>% 
  summarize(total_steps=sum(steps))
hist(steps_per_day$total_steps)
mean_steps_per_day <- mean(steps_per_day$total_steps, na.rm=TRUE)
mean_steps_per_day
```

## What is the average daily activity pattern?
Question 4: Time series plot of the average number of steps taken. 
```{r}
interval_5min <- df %>% 
  group_by(interval) %>% 
  summarize(int_steps=mean(steps, na.rm=TRUE))

plot(interval_5min$interval, interval_5min$int_steps, type="l", ylab="number of steps", xlab="interval")

```

Question 5: The 5-minute interval that, on average, contains the maximum number of steps
```{r}
max_step <- max(interval_5min$int_steps)
max_5min <- interval_5min %>% 
  filter(int_steps == max_step)
colnames(max_5min) <- c("The interval that contains the maximum number", "Maximum number of steps")
max_5min
```

## Imputing missing values
Question 6-1: Code to describe and show a strategy for imputing missing data.  The number of row which contains missing values.
```{r}
include_NA <- df %>% 
  filter(is.na(steps))
number_NA <- nrow(include_NA)
paste("The number of missing values in steps is ", number_NA, ".")
```

Question 6-2: Devise a strategy for filling in all of the missing values in the dataset -- I use the mean steps to fill the missin value.
Question 6-3: Create a new dataset -- The name of new dataset is "new_dataset".
```{r}
# I use mean values of day to fill the missing value
mean_steps <- df %>% 
  group_by(interval) %>% 
  summarize(mean_steps=mean(steps, na.rm=TRUE))

df_add_mean <- left_join(df, mean_steps, by="interval")
new_dataset <- df_add_mean %>% 
  mutate(new_steps=ifelse(is.na(steps),mean_steps,steps)) %>% 
  select(new_steps, date, interval)
```

Question 7: Make a histogram of the total number of steps in new dataset. Calculate the mean and median.
```{r}
new_steps_per_day <-  new_dataset%>% 
  group_by(date) %>% 
  summarize(total_steps=sum(new_steps))
hist(new_steps_per_day$total_steps)
new_mean_steps_per_day <- mean(new_steps_per_day$total_steps, na.rm=TRUE)
paste("The mean total number of steps taken per day is ",round(new_mean_steps_per_day), ".")
new_median_steps_per_day <- median(new_steps_per_day$total_steps, na.rm=TRUE)
paste("The mean total number of steps taken per day is ",round(new_median_steps_per_day), ".")
```

## Are there differences in activity patterns between weekdays and weekends?
Question 8-1: Create a new factor variable in the dataset with two levels – “weekday” and “weekend”
```{r}
new_dataset_weekday<- new_dataset %>% 
  mutate(weekday = wday(date)) %>% 
  mutate(weekday = ifelse(weekday %in% c(2,3,4,5,6),"weekday","weekend"))
```

Question 8-2: Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r, fig.width=8}
options(dplyr.summarise.inform = FALSE)
interval_5min_weekday <- new_dataset_weekday %>% 
  group_by(weekday,interval) %>% 
  summarize(int_steps=mean(new_steps, na.rm=TRUE))
interval_5min_weekday %>% 
  ggplot(aes(x=interval, y=int_steps, color=weekday)) +
  geom_line() +
  facet_grid(weekday ~ .) +
  theme_bw()
```
















Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
